HMM Model

### 2020-09-12 dev/raw_data_preprocessing.Rmd
Data revision, renamed vars in hmm.Rmd

```{r}
for(n in names(anonymized_data) ) {
  write_csv(anonymized_data[[n]], path = here::here("private_data", paste0(n,".csv")))
}
```

>>>>

```{r}
for(n in names(anonymized_data) ) {
  write_csv(anonymized_data[[n]], path = here::here("private_data", paste0(n,".csv")))
}
write_csv(data.frame(version = date()), path = here::here("data_revision.csv"))
```

### 2020-09-12 manuscript/hmm.Rmd
Data revision, renamed vars in hmm.Rmd

```{r}
m1_hidden_state_data <- tibble(id = factor(disease_levels, levels = disease_levels)) %>% mutate( corresponding_obs = id)

m1_observed_state_data <- tibble(id = factor(disease_levels, levels = disease_levels), is_noisy = FALSE)

@@

m1_rate_data <- rbind(
  tibble(.from = breathing_levels, .to = "Death", rate_group = "death"),
  tibble(.from = "AA", .to = "Discharged", rate_group = "improve_one"),
  tibble(.from = breathing_levels[2 : length(breathing_levels)], .to = breathing_levels[1 : (length(breathing_levels) - 1)], rate_group = "improve_one"),
  tibble(.from = breathing_levels[1 : (length(breathing_levels) - 1)], .to = breathing_levels[2 : length(breathing_levels)], rate_group = "worsen_one")
) %>%
  mutate(.rate_id = factor(1:n()),
         .from = factor(.from, levels = disease_levels),
         .to = factor(.to, levels = disease_levels)
  ) %>%
  recode_group("death") %>%
  recode_group("improve_one") %>%
  recode_group("worsen_one")

m1_initial_states <- series_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

m1_prior = brms::set_prior("normal(-2, 5)", "Intercept") +
  brms::set_prior("normal(0, 5)", "b") +
  brms::set_prior("normal(0, 2)", "sd")

m1_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group
m1 <- brmshmmdata(m1_formula, series_data, 
                  m1_rate_data, m1_hidden_state_data, m1_initial_states, prior = m1_prior,
                  observed_state_data = m1_observed_state_data)

#make_stancode_hmm(m1)
#xx <- brms::make_standata(m1$formula, data = make_data_hmm(m1)$brmsdata)
m1_fit <- brmhmm(m1, cache_file = paste0(cache_dir,"/hcq_m1.rds"))

summary(m1_fit$brmsfit)

```
```{r}
m2_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + age_norm * rate_group + sex * rate_group
m2 <- brmshmmdata(m2_formula, series_data, 
                  m1_rate_data, m1_hidden_state_data, m1_initial_states, prior = m1_prior,
                  observed_state_data = m1_observed_state_data)

m2_fit <- brmhmm(m2, cache_file = paste0(cache_dir,"/hcq_m2.rds"))

summary(m2_fit$brmsfit)
```

```{r}
m3_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + age_norm * rate_group + sex * rate_group + (rate_group | hospital_id)
m3 <- brmshmmdata(m3_formula, series_data, 
                  m1_rate_data, m1_hidden_state_data, m1_initial_states, prior = m1_prior,
                  observed_state_data = m1_observed_state_data)

m3_fit <- brmhmm(m3, cache_file = paste0(cache_dir,"/hcq_m3.rds"))

summary(m3_fit$brmsfit)
```
```{r}
m3_epred_rect <- posterior_epred_rect(m3_fit)
m3_predicted_rect <- posterior_epred_to_predicted(m3_fit, m3_epred_rect)
m3_predicted <- posterior_rect_to_long(m3_fit, m3_predicted_rect)
```

```{r}
m3_predicted_df <- posterior_long_to_df(m3_fit$data, m3_predicted)

all_ids <- m3_fit$data$serie_data %>% select(hospital_id, .serie) %>% 
  distinct() %>% arrange(hospital_id) %>% pull(.serie)
step_size <- 6
for(step in 1:ceiling(length(all_ids) / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(m3_predicted_df, all_ids[series_id], m3_fit$data) %>% print()

@@

```{r}
pp_check_last_state(m3_fit, m3_predicted_rect) #TODO separately terminal and non-terminal states. Or "state at time t" - this should be coherent
pp_check_transitions_direction(m3_fit, m3_predicted_rect)
pp_check_transitions_direction(m3_fit, m3_predicted_rect, states_from = 2:6)
```

>>>>

```{r}
hcq_m1_hidden_state_data <- tibble(id = factor(disease_levels, levels = disease_levels)) %>% mutate( corresponding_obs = id)

hcq_m1_observed_state_data <- tibble(id = factor(disease_levels, levels = disease_levels), is_noisy = FALSE)

@@

hcq_m1_rate_data <- rbind(
  tibble(.from = breathing_levels, .to = "Death", rate_group = "death"),
  tibble(.from = "AA", .to = "Discharged", rate_group = "improve_one"),
  tibble(.from = breathing_levels[2 : length(breathing_levels)], .to = breathing_levels[1 : (length(breathing_levels) - 1)], rate_group = "improve_one"),
  tibble(.from = breathing_levels[1 : (length(breathing_levels) - 1)], .to = breathing_levels[2 : length(breathing_levels)], rate_group = "worsen_one")
) %>%
  mutate(.rate_id = factor(1:n()),
         .from = factor(.from, levels = disease_levels),
         .to = factor(.to, levels = disease_levels)
  ) %>%
  recode_group("death") %>%
  recode_group("improve_one") %>%
  recode_group("worsen_one")

hcq_m1_initial_states <- series_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

hcq_m1_prior = brms::set_prior("normal(-2, 5)", "Intercept") +
  brms::set_prior("normal(0, 5)", "b") +
  brms::set_prior("normal(0, 2)", "sd")

hcq_m1_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group
hcq_m1 <- brmshmmdata(hcq_m1_formula, series_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

#make_stancode_hmm(hcq_m1)
#xx <- brms::make_standata(hcq_m1$formula, data = make_data_hmm(hcq_m1)$brmsdata)
hcq_m1_fit <- brmhmm(hcq_m1, cache_file = paste0(cache_dir,"/hcq_m1.rds"))

summary(hcq_m1_fit$brmsfit)

```
```{r}
hcq_m1_epred_rect <- posterior_epred_rect(hcq_m1_fit)
hcq_m1_predicted_rect <- posterior_epred_to_predicted(hcq_m1_fit, hcq_m1_epred_rect)
hcq_m1_predicted <- posterior_rect_to_long(hcq_m1_fit, hcq_m1_predicted_rect)
```

```{r}
hcq_m1_predicted_df <- posterior_long_to_df(hcq_m1_fit$data, hcq_m1_predicted)

all_ids <- hcq_m1_fit$data$serie_data %>% select(hospital_id, .serie) %>% 
  distinct() %>% arrange(hospital_id) %>% pull(.serie)
step_size <- 6
for(step in 1:ceiling(length(all_ids) / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(hcq_m1_predicted_df, all_ids[series_id], hcq_m1_fit$data) %>% print()

@@

```{r}
pp_check_last_state(hcq_m1_fit, hcq_m1_predicted_rect) #TODO separately terminal and non-terminal states. Or "state at time t" - this should be coherent
pp_check_transitions_direction(hcq_m1_fit, hcq_m1_predicted_rect)
pp_check_transitions_direction(hcq_m1_fit, hcq_m1_predicted_rect, states_from = 2:6)
```

```{r}
hcq_m2_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + age_norm * rate_group + sex * rate_group
hcq_m2 <- brmshmmdata(hcq_m2_formula, series_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

hcq_m2_fit <- brmhmm(hcq_m2, cache_file = paste0(cache_dir,"/hcq_m2.rds"))

summary(hcq_m2_fit$brmsfit)
```

```{r}
hcq_m2_epred_rect <- posterior_epred_rect(hcq_m2_fit)
hcq_m2_predicted_rect <- posterior_epred_to_predicted(hcq_m2_fit, hcq_m2_epred_rect)
hcq_m2_predicted <- posterior_rect_to_long(hcq_m2_fit, hcq_m2_predicted_rect)
```

```{r}
hcq_m2_predicted_df <- posterior_long_to_df(hcq_m2_fit$data, hcq_m2_predicted)

all_ids <- hcq_m2_fit$data$serie_data %>% select(hospital_id, .serie) %>% 
  distinct() %>% arrange(hospital_id) %>% pull(.serie)
step_size <- 6
for(step in 1:ceiling(length(all_ids) / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(hcq_m2_predicted_df, all_ids[series_id], hcq_m2_fit$data) %>% print()

```{r}
pp_check_last_state(hcq_m2_fit, hcq_m2_predicted_rect) #TODO separately terminal and non-terminal states. Or "state at time t" - this should be coherent
pp_check_transitions_direction(hcq_m2_fit, hcq_m2_predicted_rect)
pp_check_transitions_direction(hcq_m2_fit, hcq_m2_predicted_rect, states_from = 2:6)
```

```{r}
hcq_m3_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + age_norm * rate_group + sex * rate_group + (rate_group | hospital_id)
hcq_m3 <- brmshmmdata(hcq_m3_formula, series_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

hcq_m3_fit <- brmhmm(hcq_m3, cache_file = paste0(cache_dir,"/hcq_m3.rds"))

summary(hcq_m3_fit$brmsfit)
```
```{r}
hcq_m3_epred_rect <- posterior_epred_rect(hcq_m3_fit)
hcq_m3_predicted_rect <- posterior_epred_to_predicted(hcq_m3_fit, hcq_m3_epred_rect)
hcq_m3_predicted <- posterior_rect_to_long(hcq_m3_fit, hcq_m3_predicted_rect)
```

```{r}
hcq_m3_predicted_df <- posterior_long_to_df(hcq_m3_fit$data, hcq_m3_predicted)

all_ids <- hcq_m3_fit$data$serie_data %>% select(hospital_id, .serie) %>% 
  distinct() %>% arrange(hospital_id) %>% pull(.serie)
step_size <- 6
for(step in 1:ceiling(length(all_ids) / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(hcq_m3_predicted_df, all_ids[series_id], hcq_m3_fit$data) %>% print()

```{r}
pp_check_last_state(hcq_m3_fit, hcq_m3_predicted_rect) #TODO separately terminal and non-terminal states. Or "state at time t" - this should be coherent
pp_check_transitions_direction(hcq_m3_fit, hcq_m3_predicted_rect)
pp_check_transitions_direction(hcq_m3_fit, hcq_m3_predicted_rect, states_from = 2:6)
```

```{r}
hcq_m4_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + took_az * rate_group + age_norm * rate_group + sex * rate_group + (rate_group | hospital_id)
hcq_m4 <- brmshmmdata(hcq_m4_formula, series_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

hcq_m4_fit <- brmhmm(hcq_m4, cache_file = paste0(cache_dir,"/hcq_m4.rds"))

summary(hcq_m4_fit$brmsfit)
```

```{r}
hcq_m4_epred_rect <- posterior_epred_rect(hcq_m4_fit)
hcq_m4_predicted_rect <- posterior_epred_to_predicted(hcq_m4_fit, hcq_m4_epred_rect)
hcq_m4_predicted <- posterior_rect_to_long(hcq_m4_fit, hcq_m4_predicted_rect)
```

```{r}
hcq_m4_predicted_df <- posterior_long_to_df(hcq_m4_fit$data, hcq_m4_predicted)

all_ids <- hcq_m4_fit$data$serie_data %>% select(hospital_id, .serie) %>% 
  distinct() %>% arrange(hospital_id) %>% pull(.serie)
step_size <- 6
for(step in 1:ceiling(length(all_ids) / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(hcq_m4_predicted_df, all_ids[series_id], hcq_m4_fit$data) %>% print()

```{r}
pp_check_last_state(hcq_m4_fit, hcq_m4_predicted_rect) #TODO separately terminal and non-terminal states. Or "state at time t" - this should be coherent
pp_check_transitions_direction(hcq_m4_fit, hcq_m4_predicted_rect)
pp_check_transitions_direction(hcq_m4_fit, hcq_m4_predicted_rect, states_from = 2:6)
```

## hcq_m5 -

```{r}
hcq_m5_formula <-  ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + age_norm * rate_group + sex * rate_group + (rate_group | hospital_id) + mo(comorbidities_sum) * rate_group
hcq_m5 <- brmshmmdata(hcq_m5_formula, series_data, ☻
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

hcq_m5_fit <- brmhmm(hcq_m5, cache_file = paste0(cache_dir,"/hcq_m5.rds"))

summary(hcq_m5_fit$brmsfit)
```

```{r}
hcq_m5_epred_rect <- posterior_epred_rect(hcq_m5_fit)
hcq_m5_predicted_rect <- posterior_epred_to_predicted(hcq_m5_fit, hcq_m5_epred_rect)
hcq_m5_predicted <- posterior_rect_to_long(hcq_m5_fit, hcq_m5_predicted_rect)
```

```{r}
hcq_m5_predicted_df <- posterior_long_to_df(hcq_m5_fit$data, hcq_m5_predicted)

all_ids <- hcq_m5_fit$data$serie_data %>% select(hospital_id, .serie) %>% 
  distinct() %>% arrange(hospital_id) %>% pull(.serie)
step_size <- 6
for(step in 1:ceiling(length(all_ids) / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(hcq_m5_predicted_df, all_ids[series_id], hcq_m5_fit$data) %>% print()

```{r}
pp_check_last_state(hcq_m5_fit, hcq_m5_predicted_rect) #TODO separately terminal and non-terminal states. Or "state at time t" - this should be coherent
pp_check_transitions_direction(hcq_m5_fit, hcq_m5_predicted_rect)
pp_check_transitions_direction(hcq_m5_fit, hcq_m5_predicted_rect, states_from = 2:6)
```
